---
title: "FinalProject1"
output: pdf_document
date: "`r Sys.Date()`"
---

```{r setup, message=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# libraries
library(dplyr)
library(readxl)
library(sdcMicro)
library(readr)
?sdcMicro
set.seed(123)
```

### Loading in data from xlsx and formatting by creation of sdc object

```{r}
data <- read_excel("F/private_dataF.xlsx")
keyVars <- c("sex", "dob", "zip", "citizenship","marital_status")
sdc <- createSdcObj(dat = data, keyVars = keyVars)
```

### Generalisation and Suprresion - Top/Bottom coding to make age groups and broader regions??. K-anonimity as well!
```{r}
# generalizing `dob` by converting it to age groups
data$age <- as.numeric(format(Sys.Date(), "%Y")) - as.numeric(format(data$dob, "%Y"))
data$age <- cut(data$age, breaks = c(-Inf, 30, 50, 70, Inf), labels = c("Under 30", "30-50", "50-70", "70+"))

# generalizing `zip` into broader regions
# I dont think this one makes sense
# data$zip_region <- ifelse(data$zip %in% c(2100, 2200), "Region 1", "Region 2")

# Adding "Danish_Status" column based on citizenship
data$danish_Status <- ifelse(data$citizenship == "Denmark", "Denmark", "Non-Danish")

# updating the `sdcMicro` object with the generalized data
newKeyVars = c("sex", "age", "danish_Status", "marital_status")
sdc <- createSdcObj(dat = data, keyVars = newKeyVars)
```

### applying PRAM now to add noise to categories like party and marital_status. Randomised response will also be used on party as it is more sensitive. Is this overkill??

```{r}
# applying PRAM on `party` and `marital_status`
#sdc <- pram(sdc, variables = c("sex", "marital_status"), strata_variables = "party")
```

```{r}
# Suppressing sensitive variable values
sdc <- localSupp(sdc, threshold = 0.5, keyVar = "sex")
sdc <- localSupp(sdc, threshold = 0.5, keyVar = "marital_status")
```

### Removing the old columns
```{r}
data2 <- extractManipData(sdc)

data2 <- data2 %>% group_by(party, evote) %>% mutate(sex_new = sample(sex))
data2 <- data2 %>% group_by(party, evote) %>% mutate(marital_status_new = sample(marital_status))

# Remove columns using `dplyr::select()`
data2 <- data2 %>%
  select(-dob, -name, -citizenship, -sex, -marital_status, -zip)  # Add more columns to remove as needed

# Recreate the sdc object with the modified data
sdc <- createSdcObj(dat = data2, keyVars = c("sex_new", "age", "danish_Status", "marital_status_new"))
```

### Calculation of disclosure risks
```{r}
# Calculate and display disclosure risk
risk_value <- measure_risk(sdc, keyVars = c("sex_new", "age", "danish_Status", "marital_status_new"))
print(risk_value)
```

### Extracting the anonymized data now

```{r}
# Extract anonymized data
anonymized_data <- extractManipData(sdc)
write_csv(anonymized_data, "/Users/chris/Desktop/ITU/5thsem/SP/anonymized_dataF.csv")
```

```{r}
#anonymized_data_2 <- anonymized_data %>%
#  select(-education, -evote, -party)
#uniques <- anonymized_data_2 %>% distinct()
#uniques
```

